Cognitive and Auditory Predictors of Dysarthric Speech Understanding in Older

### 2024-01-16 Manuscript Analyses/manuscript_analysis.Rmd
reran analyses

speaker_intel <- pl_int %>%
  dplyr::mutate(type = as.factor(type),
                type = fct_relevel(type, "pretest", "posttest")) %>%
  dplyr::group_by(speaker, type) %>%
  dplyr::summarize(m_pwc = mean(pwc),
                   sd_pwc = sd(pwc)) %>%
  dplyr::ungroup() %>%
  tidyr::pivot_wider(names_from = speaker,
                     values_from = c(m_pwc:sd_pwc)) %>%
  dplyr::relocate(sd_pwc_AM1, .after = m_pwc_AM1) %>%
  gt::gt() %>%
    fmt_number(columns = "m_pwc_AM1":"sd_pwc_PDM10", decimals = 2) %>%
    tab_spanner(label = "AM1",
                columns = c(m_pwc_AM1, sd_pwc_AM1)) %>%
    tab_spanner(label = "PDM10",
                columns = c(m_pwc_PDM10, sd_pwc_PDM10)) %>%
    cols_label(type = " ",
               m_pwc_AM1 = "Mean",
               sd_pwc_AM1 = "SD",
               m_pwc_PDM10 = "Mean",
               sd_pwc_PDM10 = "SD")

@@

pl_int %>%
  ggplot() +
  aes(x = type,
      y = pwc,
      group = id) +
  geom_line(alpha = 0.3) +
  stat_summary(aes(group = speaker),
               fun.y = mean, geom = "line", linewidth = 2) +
  facet_wrap("speaker") +
  labs(x = "Phase", y = "Intelligibility") +
  theme_classic() +
  theme(legend.position = "none", aspect.ratio = 1)

@@

# Customize plot colors

library(LaCroixColoR) #remotes::install_github("johannesbjork/LaCroixColoR")
my_pal <- lacroix_palette("PeachPear", n = 15, type = "continuous")
my_pal <- rev(my_pal)

@@

Two elastic net models were conducted to evaluate the predictive relationship between the cognitive measures on pretest (initial) intelligibility: A model containing only the main effects of the cognitive measures (i.e. the main effects model) and another model that contains the main effects and all two-way interactions (i.e. the interaction model). The cross-validated R-squared and RMSE values are not significantly different between these two models. However, the r-squared value for the main effects model is higher.

The relative variable importance for the models are shown in the figure below. The most important predictor for the AM1 initial intelligibility model was words-in-noise hearing threshold, followed by working memory, the interaction between working memory and inhibitory control, and the interaction between working memory and receptive vocabulary.

@@

Plotting the interaction effects for the AM1 initial intelligibility models.

@@

effect_plot(mods1, best_win, int = F) # Words in Noise
effect_plot(mods1, list_sort_c, int = F) # Working Memory
effect_plot(mods1, list_sort_c, flanker_c) # Working Memory * Inhibitory Control
effect_plot(mods1, flanker_c, list_sort_c) # Inhibitory Control * Working Memory
effect_plot(mods1, list_sort_c, vocab_c) # Working Memory * Receptive Vocab
effect_plot(mods1, vocab_c, list_sort_c) # Receptive Vocab * Working Memory

@@

Like for AM1, two elastic net models were used to model the predictive relationship between the listener cognitive measures and pretest intelligibility for PDM10: a main effects model and an interaction model. The difference in R-square and RMSE is not significantly different between the two models. However, the R-squared value is higher for the interaction model, indicating a better fit to the data.

@@

Plotting the interaction effects for the PDM10 initial intelligibility models.

@@

effect_plot(mods2, list_sort_c, best_win) # Working Memory * Words in Noise Threshold
#effect_plot(mods2, best_win, list_sort_c) # Words in Noise Threshold * Working Memory
effect_plot(mods2, best_win, int = F) # Words in Noise Threshold
effect_plot(mods2, list_sort_c, vocab_c) # Working Memory * Receptive Vocabulary
effect_plot(mods2, vocab_c, list_sort_c)
effect_plot(mods2, card_sort_c, pattern_c) # Cognitive Flexibility * Processing Speed
effect_plot(mods2, pattern_c, card_sort_c) # Processing Speed * Cognitive Flexibility
effect_plot(mods2, flanker_c, best_win) # Inhibitory Control * Words in Noise Threshold
effect_plot(mods2, vocab_c, pattern_c) # Receptive Vocabulary * Processing Speed
effect_plot(mods2, pattern_c, vocab_c) # Processing Speed * Receptive Vocabulary
effect_plot(mods2, list_sort_c, card_sort_c) # Working Memory * Cognitive Flexibility
effect_plot(mods2, card_sort_c, list_sort_c) # Cognitive Flexibility * Working Memory

@@

Two elastic net models were used to model the relationship between the listener cognitive measures and intelligibility improvement (controlling for pretest intelligibility) for AM1: the main effects only model and the model with interactions.

@@

Plotting the interaction effects for intelligibility improvement models for AM1.

@@

effect_plot(mods12, flanker_c, best_win) # Inhibitory Control * Words in Noise Threshold
effect_plot(mods12, pattern_c, best_win) # Processing Speed * Words in Noise Threshold
effect_plot(mods12, card_sort_c, best_win) # Cognitive Flexibility * Words in Noise Threshold
effect_plot(mods12, flanker_c, vocab_c) # Inhibitory Control * Receptive Vocabulary
effect_plot(mods12, vocab_c, flanker_c) # Receptive Vocabulary * Inhibitory Control

@@

Plotting the interaction effects for the intelligibility improvement models for PDM10.

@@

# effect_plot(mods22, card_sort_c, vocab_c) # Cognitive Flexibility * Receptive Vocabulary
effect_plot(mods22, vocab_c, card_sort_c) # Receptive Vocabulary * Cognitive Flexibility
effect_plot(mods22, flanker_c, pattern_c) # Inhibitory Control * Processing Speed
effect_plot(mods22, pattern_c, flanker_c) # Processing Speed * Inhibitory Control
effect_plot(mods22, vocab_c, int = F) # Receptive Vocab 
effect_plot(mods22, flanker_c, int = F) # Inhibitory Control
effect_plot(mods22, flanker_c, best_win) # Inhibitory Control * Words in Noise Threshold
effect_plot(mods22, best_win, flanker_c) # Words in Noise Threshold * Inhibitory Control

>>>>

speaker_intel <- cor_df %>%
  dplyr::select(speaker, pretest, posttest, improvement) %>%
  tbl_summary(
    by = speaker,
    statistic = list(all_continuous() ~ "{mean} ({sd})"),
    missing = "no",
    digits = list(everything() ~ c(2)),
    label = list(pretest ~ "Pretest Inelligibility",
                   posttest ~ "Posttest Intelligibility",
                   improvement ~ "Intelligibility Change")) %>%
  as_gt()

@@

# Customize colors for figures
library(LaCroixColoR) #remotes::install_github("johannesbjork/LaCroixColoR")
my_pal <- lacroix_palette("PeachPear", type = "discrete")

pl_int %>%
  ggplot() +
  aes(x = type,
      y = pwc,
      group = id,
      color = speaker) +
  geom_line(alpha = 0.3) +
  geom_point(alpha = 0.3) +
  stat_summary(aes(group = speaker),
               fun.y = mean, geom = "line", linewidth = 2) +
  facet_wrap("speaker") +
  labs(x = "Phase", y = "Intelligibility") +
  theme_classic() +
  scale_color_manual(values = c(my_pal[5], my_pal[6])) +
  theme(legend.position = "none", aspect.ratio = 1)

@@

# Customize plot colors
my_pal <- lacroix_palette("PeachPear", n = 15, type = "continuous")
my_pal <- rev(my_pal)

@@

Two elastic net models were conducted to evaluate the predictive relationship between the cognitive measures on pretest (initial) intelligibility: A model containing only the main effects of the cognitive measures (i.e. the main effects model) and another model that contains the main effects and all two-way interactions (i.e. the interaction model). The cross-validated R-squared and RMSE values are not significantly different between these two models. However, the r-squared value for the main effects model is higher. Therefore, the main effects model is the best model for the AM1 initial intelligibility model.

The relative variable importance for the models are shown in the figure below. The most important predictor for the AM1 initial intelligibility model was words-in-noise hearing threshold, followed by the main effect of working memory.

@@

The plots below are showing the effects of words in noise hearing threshold and working memory on initial intelligibility for AM1.

-   Words in Noise Hearing Threshold has a negative relationship with initial intelligibility, indicating that those with higher words in noise thresholds have lower initial intelligibility for AM1.
-   Working Memory has a positive relationship with initial intelligibility, indicating that those listeners with greater working memory capacity had higher initial intelligibility for AM1.

@@

effect_plot(mods1, best_win, int = F) # Words in Noise
effect_plot(mods1, list_sort_c, int = F) # Working Memory

@@

Like for AM1, two elastic net models were used to model the predictive relationship between the listener cognitive measures and pretest intelligibility for PDM10: a main effects model and an interaction model. The difference in R-square and RMSE is not significantly different between the two models. However, the R-squared value is higher for the interaction model and multiple interactions were indicated as important predictors.

@@

The figures below are plotting the interaction effects for the PDM10 initial intelligibility models.

-   The interaction between working memory and words in noise hearing threshold indicates that working memory did not have much of an effect for listeners with lower (i.e. better) hearing thresholds on initial intelligibility for PDM10. However, working memory help compensate for listeners with elevated hearing thresholds.
-   Like with AM1, words in noise hearing threshold has a negative relationship with initial intelligibility for PDM10. 
-   The interaction between working memory and receptive vocabulary indicates that greater receptive vocabulary knowledge helped compensate for listeners with lower working memory capacity. 
-   The interaction between cognitive flexibility and processing speech indicates that greater cognitive flexibility helped compensate for listener with lower processing speed. However, greater processing speed tends to have a negative effect on initial intelligibility for listeners with greater cognitive flexibility.
-   The interaction between inhibitory control and words in noise hearing threshold indicate that inhibitory control helped compensate for listeners with slightly elevated words in noise hearing thresholds. However, inhibitory control did not have a great effect on initial intelligibility for listeners with low or elevated hearing thresholds.
-   The interaction between receptive vocabulary and processing speed indicates that those with greater processing speed and greater vocabulary had greater initial intelligibility
-   The interaction between cognitive flexibility and working memory indicates that cognitive flexibility helps compensate perception for listeners with lower working memory capacity, but may not have much of an effect (or even a possible negative effect) for listeners with higher working memory capacity.

@@

effect_plot(mods2, list_sort_c, best_win) # Working Memory * Words in Noise Threshold
effect_plot(mods2, best_win, int = F) # Words in Noise Threshold
effect_plot(mods2, vocab_c, list_sort_c) # Receptive Vocabulary * Working Memory
effect_plot(mods2, card_sort_c, pattern_c) # Cognitive Flexibility * Processing Speed
effect_plot(mods2, pattern_c, card_sort_c) # Processing Speed * Cognitive Flexibility
effect_plot(mods2, flanker_c, best_win) # Inhibitory Control * Words in Noise Threshold
effect_plot(mods2, best_win, flanker_c) # Words in Noise Threshold * Inhibitory Control
effect_plot(mods2, vocab_c, pattern_c) # Receptive Vocabulary * Processing Speed
effect_plot(mods2, pattern_c, vocab_c) # Processing Speed * Receptive Vocabulary
effect_plot(mods2, list_sort_c, card_sort_c) # Working Memory * Cognitive Flexibility
effect_plot(mods2, card_sort_c, list_sort_c) # Cognitive Flexibility * Working Memory

@@

Two elastic net models were used to model the relationship between the listener cognitive measures and intelligibility improvement (controlling for pretest intelligibility) for AM1: the main effects only model and the model with interactions. The difference in RMSE and R-squared is not significantly different. However, multiple interactions were indicated to be important predictors in the model.

@@

The the interaction effects for the intelligibility improvement model for AM1 are plotted below.

-   The interaction between inhibitory control and words in noise threshold indicate that the effect of inhibiotry control on intelligibility improvement was reduced for listeners with higher hearing thresholds
-   The interaction between processing speed and words in noise threshold indicate that greater processing speed reduced the effect on intelligibility improvement for listeners with higher thresholds.***
- The interaction between cognitive flexibility and words in noise threshold indicate that greater cognitive flexibility helped compensate for intelligibility improvement for listeners with high thresholds. However, for listeners with lower thresholds, greater cognitive flexibility had a negative effect on improvement.
-   The interaction between inhibitory control and receptive vocabulary indicates that increased vocabulary knowledge had a positive effect on intelligibility improvement for listeners with low and high inhibitory control. However, vocabulary knowledge had a negative effect for listeners with "mid/average" inhibitory control.

@@

effect_plot(mods12, flanker_c, best_win) # Inhibitory Control * Words in Noise Threshold
effect_plot(mods12, best_win, flanker_c) # Words in Noise * Inhibitory Control
effect_plot(mods12, pattern_c, best_win, num = 4) # Processing Speed * Words in Noise Threshold
effect_plot(mods12, best_win, pattern_c, num = 4)
effect_plot(mods12, card_sort_c, best_win, num = 5) # Cognitive Flexibility * Words in Noise Threshold
effect_plot(mods12, flanker_c, vocab_c, num = 5) # Inhibitory Control * Receptive Vocabulary
effect_plot(mods12, vocab_c, flanker_c, num = 5) # Receptive Vocabulary * Inhibitory Control

@@

The interaction effects for the intelligibility improvement models for PDM10 are plotted below.

-   The interaction between cognitive flexibility and receptive vocabulary indicates that cognitive flexibility had a positive effect on improvement for listeners with higher vocabulary knowledge, but minimal effect for listeners with lower vocabulary knowledge
-   The interaction effect between inhibitory control and processing speed indicates that processing speed has a negative effect on intelligibility improvement for listeners with mid-range inhibitory control. However, processing speed has minimal effect on improvement when listeners have lower or higher inhibitory control scores.***
-   Receptive vocabulary has a positive effect on intelligibility improvement.
-   Inhibitory control has a negative effect on intelligibility improvement.
-   The interaction between inhibitory control and words in noise threshold indicates a positive effect of inhibitory control for listeners with low words in noise thresholds but little to no effect for listeners with elevated hearing thresholds (a potential negative effect)

@@

effect_plot(mods22, card_sort_c, vocab_c) # Cognitive Flexibility * Receptive Vocabulary
effect_plot(mods22, vocab_c, card_sort_c) # Receptive Vocabulary * Cognitive Flexibility
effect_plot(mods22, flanker_c, pattern_c) # Inhibitory Control * Processing Speed
effect_plot(mods22, pattern_c, flanker_c) # Processing Speed * Inhibitory Control
effect_plot(mods22, vocab_c, int = F) # Receptive Vocab 
effect_plot(mods22, flanker_c, int = F) # Inhibitory Control
effect_plot(mods22, flanker_c, best_win, num = 4) # Inhibitory Control * Words in Noise Threshold
